\documentclass{ximera}

\input{preamble}


\title{All In The Timing}
\author{Vic Ferdinand, Betsy McNeal, Robin Pemantle, Jenny Sheldon}

\begin{document}
\begin{abstract} \end{abstract}
\maketitle



\begin{problem}
 You take your fifth grade class to the local carnival, where there are three games to rip off even the best of players.  These games cost 21 cents, 8 cents, and 18 cents, respectively, to play.  Your most pesky student, Gullible Gil, forgot to bring money from home and hounds you to lend some to him.  You lend him some and he gets addicted to the first game.  He plays until he's out of money, but then wants to play the second game.  To shut him up, you give him the same amount as before.  Gil comes back broke again - begging and whining to play the third game.  Once again, you give the same amount to Gil.  Of course, he blows all of this money at the third game with no prize for you.  How much money will you demand from Gil's mother, No Pay Pauline?
\end{problem}

\begin{problem}
 While trying to avoid Gil, you sneak in behind one of the slot machines and rig it to come up with three cherries (one in each window to hit the jackpot) at a time known only to you. You know the first window will come up with a cherry every 15 times, the second window every 24 times, and the third window every 40 times.  How many times should the machine be played before you step in to win your ill-gotten gain of $\$$1.98?

\end{problem}

\newpage
\begin{instructorNotes}
This is usually the first activity we do towards introducing the Least Common Multiple and Greatest Common Factor.  For this reason, students are not necessarily expecting to find either of these values when they first approach the problems.  Note that each problem has several answers, so students will not necessarily all have the same correct answer for this problem. We use this observation to highlight that sometimes we want the least common multiple, and sometimes we want just any common multiple.

\begin{itemize}
	\item In our experience, most students come up with the idea of the LCM from listing by ``brute force''. This is a good starting method for the children they will teach, so we make a point to discuss it.
	\item If time, we sometimes extend the discussion by asking how their methods would change if the numbers were ``not so nice''.  This can push the discussion towards prime factorization if no students have suggested that method already.  When we do this activity after talking about prime factorization, some students will usually have the idea to factor these numbers.
	\item We like this introduction to LCM because it helps us make the point that the LCM is a ``common time'' of things that grow at a constant rate. %(``slope with $y$-intercept = 0'')
\end{itemize}

\timing{We usually use a half of a class period for this activity.  We give students about 10 minutes to think about and solve the problems, then discuss for 20 minutes or so.  We then will move on to another activity, or take the remaining time to introduce the ideas of GCF and LCM.}
\end{instructorNotes}




\end{document}