\documentclass[handout]{ximera}
\input{preamble}

\title{Back And Forth}
\author{Vic Ferdinand, Betsy McNeal, Jenny Sheldon}

\begin{document}
\begin{abstract}\end{abstract}
\maketitle

\begin{instructorIntro}
This activity is an introduction to the concept of measurement conversion.  We will study the topic of measurement conversion throughout the semester, as we consider area and volume in more depth.  Most of the examples treated at this stage should concern length.

The activity is designed to bring up some misconceptions about measurement conversion, as well as give students a chance to do conversions without their preconceived notions.  The idea of ``dimensional analysis'', or making a fence-post chart, should be avoided.  Students should use reasoning, and be prepared to explain their answers in terms of groups and objects per group, as discussed in Math 1125.

\timing{This activity should take about half a class period.  Give the students a few minutes to think about the first problem, discuss, and then move on to the second problem.}

\end{instructorIntro}


\begin{problem}
    Jack and Jill are measuring the lengths of their desks.  The desks are identical, but Jack is using a piece of pencil lead to measure the length and Jill is using a necklace that she stretched out into a straight line.  Jack reports that his desk measures 20 units while Jill reports that her desk measures 5 units.
    \begin{enumerate}
        \item Jack claims that his desk must be bigger than Jill's, since his desk is 20 units long.  Is Jack correct?  Why or why not?
        \item Jill claims that her desk must be bigger than Jack's, since she used a bigger unit to measure the desk.  Is Jill correct?  Why or why not?
        \item What is the relationship between the two units used to measure the desks?  Justify your answer.
        \item Next, Jill measures the length of the chalkboard and finds it to be 13.5 units long.  When Jack measures the chalkboard, what will his answer be?  Justify your answer.  (Hint: some Math 1125 terminology should make its way into your explanation!)
    \end{enumerate}
    
    \begin{solution}
    \begin{enumerate}
    \item Jack is not correct: the desks are identical.  Using a different unit can result in a different number for the length of the desk, even though the desks didn't change.
    \item Jill is not correct: the desks are identical.  A larger unit just results in a smaller number of total units used, even thought the desks didn't change.
    \item The relationship is multiplicative: there are four of Jack's unit for every one of Jill's.
    \item Since there are four of Jack's unit for every one of Jill's, the chalkboard measures $13.5 \times 4 = 54$ of Jack's unit.
    \end{enumerate}
    \end{solution}
    
    \begin{instructorNotes}
    Here students confront two basic misconceptions about measurement conversion, and then convert using non-standard units of length.  The misconceptions are common amongst children - even though the desks are identical, getting different numbers for the measurement can be confusing to children.  Have students explain the misconceptions clearly for the class.  The question about the relationship between the units is intentionally vague.  Students can relate this process to several topics covered in Math 1125, including multiplication and ratios.
    \end{instructorNotes}
\end{problem}

\begin{problem}
    Jason knows that there are 3 feet in one yard.  He also knows that a football field is 100 yards long.  Since a yard is bigger than a foot, Jason thinks that a football field should be $100 \div 3 = 33 \frac{1}{3}$ feet long.
    \begin{enumerate}
    \item Explain Jason's reasoning.  What does he think we should do, and why did he come to that conclusion?
    \item What is the correct answer to this problem? Explain.
    \item How would you help Jason to understand his mistake?  Give at least two methods you could try.
    \end{enumerate}
    
    \begin{solution}
        \begin{enumerate}
        \item Jason thinks we should divide by three since there are three feet in one yard.  Since feet are smaller than yards, and division (sometimes) makes things smaller, he reasons we should divide.
        \item We should multiply by three in order to find the correct answer: $100 \times 3 = 300$ feet.
        \item To help Jason understand his mistake, we could draw a picture of a football field, and divide each yard into feet.  We could then count the number of feet total to find the correct answer.  We could also return to the definition of multiplication and division, and ask Jason to identify groups and objects per group in this situation.
        \end{enumerate}
    \end{solution}
    
    \begin{instructorNotes}
    Answers will probably vary for students' ideas about Jason's reasoning.  One important underlying misconception to discuss is that of ``multiplication makes things larger, and division makes things smaller''.  This will hopefully come up naturally in discussion, as it is one reason that Jason might come to this conclusion.  
    
    Students should try to find as many ways as they can to help Jason with his problem - including drawing accurate pictures (perhaps with a smaller number of yards).  Drawing good pictures is, as usual, a very age-appropriate task for grades K-3.
    \end{instructorNotes}
\end{problem}


\begin{problem}
These problems remind Mike of fraction comparison.  What might he be thinking?

\begin{instructorNotes}
This is a challenge/bonus question!  The idea is that more pieces in the denominator results in you needing more of those pieces to make the same amount.
\end{instructorNotes}
\end{problem}


\end{document}