\documentclass{ximera}

\input{preamble}


\title{Divisibility With Remainders}
\author{Vic Ferdinand, Betsy McNeal, Jenny Sheldon}

\begin{document}
\begin{abstract} \end{abstract}
\maketitle

\begin{instructorIntro}
The idea of this activity is to engage students in discussion of more general conclusions that can be made about divisibility relationships.  This activity is also nice for exploring conjectures about some numbers or a class of numbers versus conjectures about all numbers.  We guide students through the ``observe, conjecture, justify'' process.

\begin{itemize}
    \item For each of the problems, students should be encouraged to justify their answers in more than one way.  It is good to have an algebraic method as well as a pictorial method demonstrated for the class.
    \item Students should begin by working with concrete examples, but then extend to a general argument.
\end{itemize}

{\bf Suggested Timing:} This activity can take between a half-class and a full class, depending on the time you have.  For a half-class, just focus on the first problem, giving students 5-10 minutes to think about the problem, and then 15-20 minutes to discuss.  In that case, you will want to try to generalize as a whole class.  For a full class, repeat this again with the second case, and then give students time to try to generalize on their own before you discuss.

\end{instructorIntro}

\begin{problem} Let's use $A$ to represent a number that has a remainder of 1 when you divide by $5$, and let's use $B$ to represent a number that has a remainder of 4 when you divide by $5$.
\begin{enumerate}
\item Write down some possibilities for the numbers $A$ and $B$.
\item  What is the remainder when you calculate $(A + B) \div 5$?  Start by working with your examples, then make an argument about what should be true for any $A$ and $B$.
\item  What is the remainder when you calculate $(AB) \div 5$?  Start by working with your examples, then make an argument about what should be true for any $A$ and $B$.
\end{enumerate}
\end{problem}

\begin{problem}
Now let's use $C$ to represent a number that has a remainder of 2 when you divide by $5$.
\begin{enumerate}
\item  Write down some possibilities for the numbers $C$.
\item   What is the remainder when you calculate $(C + B) \div 5$?  
\item  What is the remainder when you calculate $(CB) \div 5$?
\end{enumerate}
\end{problem}

\begin{problem}
  Is there a general rule, here?  If so, what is it?  Describe it as accurately as you can, and explain why your rule should always hold.
\end{problem}


 



\end{document}