\documentclass{ximera}

\input{preamble}


\title{Prime Numbers From $1$ to $120$}
\author{Vic Ferdinand, Betsy McNeal, Jenny Sheldon}

\begin{document}
\begin{abstract} \end{abstract}
\maketitle

\begin{instructorIntro}
The goal of this activity is to introduce the prime numbers, and to discuss the method called the ``Sieve of Eratosthenes'' for finding prime numbers.

Also, we can use this activity to discuss quick methods for determining whether a number is prime - checking only prime factors, and checking up to the square root of the number - if you have time for the final problem.  Make sure to have students thoroughly explain the use of each of these ideas.

{\bf Suggested Timing:} Give students approximately 5-10 minutes to come up with ideas about how to systematically find prime numbers, then spend about 15 minutes having them describe their methods and discussing whether they could help find all prime numbers.  Then give students another 5-10 minutes to think about the final problem, and 15 minutes for discussion.
\end{instructorIntro}

\begin{problem}
Although you may know many of the prime numbers from 1 to 120, figure
out a way to find them systematically  \textit{without dividing}!

\vspace*{1cm}

\begin{tabular}{r r r r r r r r r r}

  1 &   2 &   3 &   4 &   5 &   6 &   7 &   8 &   9 &  10\\
  \\
 11 &  12 &  13 &  14 &  15 &  16 &  17 &  18 &  19 &  20\\
 \\
 21 &  22 &  23 &  24 &  25 &  26 &  27 &  28 &  29 &  30\\
 \\
 31 &  32 &  33 &  34 &  35 &  36 &  37 &  38 &  39 &  40\\
 \\
 41 &  42 &  43 &  44 &  45 &  46 &  47 &  48 &  49 &  50\\
 \\
 51 &  52 &  53 &  54 &  55 &  56 &  57 &  58 &  59 &  60\\
 \\
 61 &  62 &  63 &  64 &  65 &  66 &  67 &  68 &  69 &  70\\
 \\
 71 &  72 &  73 &  74 &  75 &  76 &  77 &  78 &  79 &  80\\
 \\
 81 &  82 &  83 &  84 &  85 &  86 &  87 &  88 &  89 &  90\\
 \\
 91 &  92 &  93 &  94 &  95 &  96 &  97 &  98 &  99 & 100\\
 \\
101 & 102 & 103 & 104 & 105 & 106 & 107 & 108 & 109 & 110\\
\\
111 & 112 & 113 & 114 & 115 & 116 & 117 & 118 & 119 & 120\\
\end{tabular}

\end{problem}
\begin{problem}
 Could your method work to find all of the primes to 320?
\end{problem}

%\item DELETE? Find all of the prime factors of 1008. How can you be sure you've found them all?


\begin{problem}
Consider the numbers 239, 323, and 4001. Are any of these prime? How
can you be sure?
\end{problem}



\end{document}