\documentclass[nooutcomes]{ximera}
\input{preamble}
\title{Return to Walk the Line}
\author{Vic Ferdinand, Betsy McNeal, Jenny Sheldon}

\begin{document}
\begin{abstract}
 (You have seen some of these problems earlier this semester.)
\end{abstract}
\maketitle



\begin{problem}
 Free-Lance Freddy works for varying hourly rates, depending on the
job.  He also carries some spare cash for lunch.  To make his
customers sweat, Freddy keeps a meter on his belt telling how much
money they currently owe (with his lunch money added in).  On Monday, 3 hours into his work as a gourmet burger flipper,
  Freddy's meter reads $\$42$. 7 hours into his work, his meter reads
  $\$86$.  
  \begin{enumerate}
  \item Draw a graph that could represent Freddy's situation. Be ready to explain how you drew your graph and how each part of your graph represents the story situation.
  \item Earlier this semester you answered the following question using a table:  {\bf If he works for 9 hours, how much money will he have?  When
  will he have $\$75$?}  Can you use your graph to answer the same question?
  \end{enumerate}
\end{problem}


\begin{problem}
\emph{Slimy Sam is on the lam from the law.  Being not-too-smart, he drives
the clunker of a car he stole east on I-70 across Ohio.  Because the
car can only go a maximum of 52 miles per hour, he floors it all the
way from where he stole the car (just now at the Rest Area 5 miles
west of the Indiana line) and goes as far as he can before running out
of gas 3.78 hours from now.}
 \begin{enumerate}
  \item Draw a graph that could represent Slimy's situation.  Be ready to explain how you drew your graph and how each part of your graph represents the story situation.

  \item Earlier this semester you answered the following questions using a table:  
  \begin{enumerate}[label=(\roman*)]
   {\bf \item Where will he be 3 hours after stealing the car?
 \item  Where will he be when he runs out of gas and is arrested?
 \item When will he be at mile marker 99 (east of Indiana)?
\item When will he be at mile marker 71.84?}
\end{enumerate}
Can you use your graph to answer the same questions?

\end{enumerate}

\end{problem}

\newpage
\begin{problem} 
\emph{Counterfeit Cathy sells two kinds of fake cereal: Square Cheerios for
$\$$4 per pound and Sugarless Sugar Pops for $\$$5 per pound.  If Cathy's goal for today is to sell $\$$1000 of cereal, how
  much of each kind should she sell?}

\begin{enumerate}
  \item Can you answer this question using a table? If yes, be ready to describe how you used the table to answer this question. If no, explain why not.
    \item Draw a graph that could represent Cathy's situation.  Be ready to explain how you drew your graph and how each part of your graph represents the story situation.
  \item Can you use your graph to answer the same questions?
  \end{enumerate}

\end{problem}

\begin{problem}
For each of the situations above, write an equation relating the two
variables (hours and current financial status, hours and position, 
pounds of Square Cheerios and pounds of Sugarless Sugar Pops) and
answer the following questions:
\begin{enumerate}
\item Could you have used your equations to help you solve the problems
  above? Are all three equations similar to each other? If yes, how are they alike?  If no, how are they different?
   \item How does each equation reflect the table you made earlier in the semester?
  \item  What are the different features of your graphs represented in each equation?
 \end{enumerate}

\end{problem}

\begin{problem} \label{Return5}
Compare and contrast the graphs you have drawn in this activity.  Make conjectures about why your observations are true.

\end{problem}

\begin{problem} 
These three problems all involve scaling.  Looking back over your work, what is the role of scaling in these three problems?


\end{problem}


\newpage
\begin{instructorNotes}
The goal of this activity is to understand the connection between linear relationships, slope, and scaling.  These ideas will be represented in both equations and graphs.  This activity uses the problems from the ``Walk the Line'' activity, and refers to students having already completed these problems.  Where the original ``Walk the Line'' activity has problems that refer to the point-slope as well as slope-intercept form of a line, the current activity adds the general form of a line as well.  The original activity was mostly intended to relate lines and arithmetic sequences, so this version's emphasis on graphs and equations is also different from the original.

On our calendar, this activity is the first we do on the topic of similarity and scaling.  No prior familiarity with any of these ideas is assumed. 




 \begin{itemize}
    \item The first two problems here go back to the problems from the original Walk the Line and make sure that students answer those questions using graphs and equations if they have not already done so.  Our goal is to focus on the connection between these equations and their tables and graphs. Once we have established the connection between constant rate of change and linear graphs and equations, then we discuss the relationship with scaling.
    \item In Problem \ref{Return5}, we would like the students to connect the idea of constant rate of change with constant slope.  In particular, they should recognize that all these situations have lines as their graphs, and that lines represent a constant rate of change or constant slope. We discuss this idea from ``both directions'': if a graph has constant slope, must it be a line?  If a graph is a line, must it have constant slope?
    \item Our goal in the final problem is to help students equate ``slope" with ``same ratio" with ``scaling".  This might take the form of observing similar triangles when they sketch the ``over and up", or perhaps in the verbal formulation that ``increasing the $x$ value by a factor 3 will result an increase (or decrease) of the $y$ value by a factor of 3".  We have another activity on this topic called ``Orange You Glad We're Back to Ratios''.
\end{itemize}



\timing{We usually allot one class period for this activity, but fully discussing all the problems would take more time than that.  Generally, we give students 10 minutes to work through the first two problems which refer back to the original activity, then discuss for 20 minutes.  We typically skip the general form of a line.  We then give students 5 more minutes to compare and contrast and think about scaling, and use the rest of the time for discussion on these new ideas.}
\end{instructorNotes}






\end{document}