\documentclass[nooutcomes]{ximera}

\input{preamble}


\title{Searching Lockers}
\author{Vic Ferdinand, Betsy McNeal, Robin Pemantle, Jenny Sheldon}

\begin{document}
\begin{abstract} \end{abstract}
\maketitle


\begin{problem}
At Fermat High School, where there are 100,000 lockers and 100,000 students, a strange tradition has developed over the years.  Every April 1, the students line up outside of the building and go into the building one at a time.  Student $\#1$ starts at Locker $\#1$ and opens every locker.  Then, Student $\#2$ starts at Locker $\#2$  and closes every other locker.  Then Student $\#3$ starts at Locker $\#3$ and ``switches" every third locker (i.e., closes it if it was open and opens it if it was closed).  Then Student $\#4$ switches every fourth locker.  If this process continued, after all the students had taken their turns, which lockers would be open?  Why?
\end{problem}
\newpage

\begin{instructorNotes}
This is often the first number theory activity that we do.  Solving the problem gives us the chance to define some terminology related to number theory, such as even, odd, prime, factor, multiple, and so on, as the students use these ideas to explain their thinking about the problem.

The activity is also a chance for students to do some problem solving - it's unusual for someone to see the pattern right away.  We usually see students trying to look up the problem on the internet (with varying levels of success), or trying to guess and check as a first strategy.  When students are stuck, we often suggest that they try a smaller number of lockers to get started, though many students will try 10 lockers to start and not be convinced that they found much of a pattern.  A larger number than $10$ tends to be more illuminating.

Another strategy we have found helpful is to stop students every few minutes to discuss as a whole class the kinds of strategies they are using.  We also eventually start asking the students questions like, ``How many students touch Locker 5?'', ``Which students will touch Locker 8? Locker 12?  Locker 13?'', or ``Is there a point after which Locker 23 is not touched again?''

{\bf Suggested Timing:} This takes us about a full class period, but we have also had success with splitting the activity over two periods so that students can go home and think about some of these ideas without immediate resolution.  In that case, we use about half a class each day.
\end{instructorNotes}

\end{document}