\documentclass{ximera}

\input{preamble}


\title{Statements About Factors}
\author{Vic Ferdinand, Betsy McNeal, Jenny Sheldon}

\begin{document}
\begin{abstract} \end{abstract}
\maketitle




Remember, \emph{a is a factor of b} if there is some integer $q$ such that $b = a \cdot q$.
\begin{problem}
 Make some true statements using this terminology.
\end{problem}

\begin{problem}
Use the definition of \emph{factor} to decide which of the following are true and which are false. If it is true, find $q$ satisfying the definition of a factor. If it is false, give an explanation.

\begin{enumerate}
\item 7 is a factor of 56 
\item \label{Statementsb} 21 is a factor of 2121
\item \label{Statementsc} 3 is a factor of $9 \times 41$
\item \label{Statementsd} 2 is a factor of $(3^{16} \times 5^4 \times 7^8)$
\item \label{Statementse} 6 is a factor of $(2^4 \times 3^2 \times 7^3 \times 13^5)$?
\item \label{Statementsf} 100000 is a factor of  $(2^3 \times 3^9 \times 5^{11} \times17^8)$
\item \label{Statementsg} 6000  is a factor of  $(2^{21} \times 3^{17} \times 5^{89} \times 29^{37})$
\item \label{Statementsh} $p^3q^5r$  is a factor of $(p^5q^{13}r^7s^2t^{27}) $
\item \label{Statementsi} 7 is a factor of $(5 \times (21+14))$
\end{enumerate}
\end{problem}


\newpage
\begin{instructorNotes}
The purpose of this activity is to give students more practice with the definition of factors, and more practice with exponential notation.

\begin{itemize}
    \item If you haven't yet, you will likely want to review exponential notation.
    \item This activity can make some later activities about GCF and LCM a little easier to handle.
\item We often don't have time for all of these problems, so you can pick a subset which seems to cover most of the ideas, like \ref{Statementsb}, \ref{Statementse}, \ref{Statementsf}, and \ref{Statementsh}.
\item A related question to \ref{Statementsf} is ``How many zeros are at the end of the number $2^3\times 3^9\times 5^{11} \times 17^8$?''.  This can later make some nice connections with the ``Shampoo'' activity.
\end{itemize}







{\bf Suggested Timing:} If you fully explore all parts of this activity, it will take a full class period.  To fit it in half, choose some parts and leave others for homework.
\end{instructorNotes}

\end{document}