\documentclass[nooutcomes]{ximera}
\input{preamble}

\title{Where Are We?}
\author{Vic Ferdinand, Betsy McNeal, Jenny Sheldon}
\begin{document}
\begin{abstract}\end{abstract}
\maketitle


\begin{problem}
Assume you can only stand at some place along one wall of a room.  How could you describe where you are using numbers?  What is the least information that would communicate your location?

\end{problem}

\begin{problem}
Now assume you can stand anywhere on the floor in the room.  How could you describe where you are using numbers?  What is the least information that would communicate your location?


\end{problem}



\begin{problem}
Describe the location of at least 5 points on the floor using your system from number 2.
\end{problem}


\begin{problem}\label{WhereAreWe4}
Assume you have a grid laid out in the Cartesian system.  Describe and draw ALL locations where:
\begin{enumerate}
\item The y-coordinate is 5.
\item The x-coordinate is -3.
\item The sum of the coordinates is 10.
\item The product of the coordinates is 24.
\item The y-coordinate is the same as the x-coordinate.
\item The y-coordinate is double the x-coordinate.
\item The y-coordinate is 3 less than double the x-coordinate.
\item The x-coordinate is the square of the y-coordinate.
\end{enumerate}

\end{problem}

\begin{problem}
Now try to describe any point in the air in the room in as many ways as possible.
\end{problem}

\begin{problem}
Let's return to the earlier activity, I Walk the Line.
    \begin{enumerate}
        \item Graph your solutions to at least two of the problems.  Does it make sense to ``connect the dots'' on each graph?
        \item Write an equation that describes each story in terms of the coordinates that you used.
        \item How are your equations and graphs alike? How are they different?
    \end{enumerate}    

\end{problem}

\newpage
\begin{instructorNotes}
Here are our goals for this activity.
\begin{enumerate}
\item We want students to think about our usual coordinate system from a new perspective, i.e., beginning from the need to locate something in the plane or in space.  This is an instance of ''making the familiar strange'', as we are asking students to work from a perspective they have not previously considered.
\item  We want students to think about the essential features of any location system: reference point, direction, unit of distance.
\item We want students to describe the locations in this activity's fourth problem using explanations of the meaning of coordinates, rather than through memorized formulae.
\end{enumerate}

In our calendar, this activity is the first of two we do to introduce the coordinate plane and graphing.  This activity follows ``I walk the Line'' about linear relationships, and is followed by ``The Graphic Details'' where we discuss the significance of different parts of a graph.

\begin{itemize}
\item  We have observed that students only need to work through about four or five of the parts in the ``Describe and draw all locations'' problem.  We sometimes assign the remaining questions for homework.
\item  The last question is intended to connect this activity's ideas to the work done in ``I Walk the Line''.  Here, we do want students to think about the independent and dependent variables, plot these to provide a second representation of the charts done earlier, and finally find an equation for each situation as a third representation of the same relationship.  In ``I Walk the Line'', we were much more focused on representing the relationship algebraically.
\end{itemize}



\timing{This activity is allotted a full class period in our calendar.  We give students 5-10 minutes to work in small groups on the first three problems, then discuss their solutions and the essential features of a coordinate system.  We then choose a subset of Problem \ref{WhereAreWe4} for the students to work on, and give about 5 minutes in small groups.  After discussing their solutions to these problems, we use the rest of the time (if any) to discuss the final problems.}
\end{instructorNotes}






\end{document}