\documentclass{ximera}
\input{preamble}
\title{You Can Count On It}
\author{Vic Ferdinand, Betsy McNeal, Robin Pemantle, Jenny Sheldon}

\begin{document}
\begin{abstract}
\end{abstract}
\maketitle

\begin{instructorIntro}
This activity is intended to introduce our counting activities.  Special attention should be paid to applications of multiplication in counting, and the way that representations of multiplication - arrays, ordered lists, tree diagrams - are used.
\end{instructorIntro}

\begin{problem}
Nikki has three different stores to visit this afternoon: Target, WalMart, and Meijer.  How many ways are there for Nikki to complete her shopping trip?  Explain.
\end{problem}

%\begin{problem}
%Calvin is trying to complete the ``Columbus Coffee Experience", which includes buying coffee at five different locations.  

%\begin{enumerate}
%\item Assuming Calvin chooses five locations in advance, how many ways are there for him to reach his goal?
%\item Assuming Calvin does not choose the locations in advance, and there are 10 participating coffee shops, how many ways are there for him to reach his goal?
%\end{enumerate}
%\end{problem}



\begin{problem}
A certain passcode must be made in the following way: choose one of O, S ,or U, then choose a whole number between 6 and 9 (inclusive).  How many passcodes can be made?  Answer this question in the following ways:
\begin{enumerate}
\item By writing down all of the possibilities.
\item By organizing the possibilities in a table.
\item By drawing a tree diagram containing all of the possibilities.
\end{enumerate}
\end{problem}

\begin{problem}
You are a contestant on ``The Amazing Race''.  You must travel from your home to New York City, then from New York City to London, and finally from London to Paris.  You will travel from your home to New York City by car, train, bus, or plane.  You will travel from New York City to London by ship or by plane.  Finally, you will travel from London to Paris by ship, plane, or helicopter.
\begin{enumerate}
\item How many trips are possible from home to New York City to London?
\item How many trips are possible from home to New York City to London to Paris?
\end{enumerate}
\end{problem}

\begin{problem}
From 2001 - 2004, an Ohio license plate consisted of two letters followed by two digits followed by two letters.
\begin{enumerate}
\item How many different Ohio license plates could be made?
\item How many different Ohio license plates could be made if there are no repeats of numbers or letters allowed?
\end{enumerate}
\end{problem}



\end{document}