%\documentclass[handout]{ximera}
\documentclass{ximera}

\input{../preamble}

\title{Divisibility Statements}
\author{Bart Snapp and Brad Findell}

\outcome{Learning outcome goes here.}

\begin{document}
\begin{abstract}
Abstract goes here.  
\end{abstract}
\maketitle

\label{A:divisibilityStatements}

Let $a|b$ mean $b=aq$ for some integer $q$.  (Read $a|b$ as ``$a$ divides $b$''.)  

\begin{problem}
Using the numbers 56 and 7, make some true statements using the notation above and one or more of the words factor, multiple, divisor, and divides.  
\end{problem}

\begin{problem}
Use the definition of \emph{divides} to decide which of the following are true and which are false.  If a statement is true, find $q$ satisfying the definition of divides.  If it is false, give an explanation.  (Hint:  Try to reason about multiplication without actually multiplying.)
\begin{enumerate}
\item $21|2121$
\item $3|(9\times 41)$
\item $6|(2^4\times 3^2\times 7^3\times 13^5)$
\item $100000|(2^3\times 3^9\times 5^{11}\times 17^8)$
\item $6000|(2^{21}\times 3^7 \times 5^{17}\times 29^5)$
\item $p^3q^5r|(p^5q^{13}r^7s^2t^{27})$
\item $7|(5\times 21 + 14)$
\end{enumerate}
\end{problem}

\begin{teachingnote}
Some students will write $2^4\cdot 3^2=6^6$.  Some students will insist on multiplying and dividing with their calculators. 

The following problems are challenging but worth the time.
\end{teachingnote}

\begin{problem}
If $a|b$ does $a|(bc)$?  Explain. 
\end{problem}

\begin{problem}
If $a|(bc)$ does $a|b$?  Explain. 
\end{problem}

\begin{problem}
If $a|b$ and $a|c$ does $a|(b+c)$?  Explain.  
\end{problem}

\begin{problem}
If $a|(b+c)$ does $a|b$ and $a|c$?  Explain.  
\end{problem}

\begin{problem}
If $a|(b+c)$ and $a|c$ does $a|b$?  Explain.  
\end{problem}

\begin{problem}
Suppose that $$(3^5\cdot 7^9\cdot 11^x\cdot 13^y)|(3^a\cdot 7^b\cdot 11^{19}\cdot 13^7)$$
What values of $a$, $b$, $x$, and $y$ make true statements? 
\end{problem}

\end{document}
