%\documentclass[handout]{ximera}
\documentclass[nooutcomes]{ximera}

\input{../preamble}

\title{What Can Division Mean?}
\author{Bart Snapp and Brad Findell}

\outcome{Learning outcome goes here.}

\begin{document}
\begin{abstract}
We explore meaning of division.
\end{abstract}
\maketitle

\label{A:dm}
\begin{teachingnote}
Two models of division:  
\begin{itemize}
\item ``How many in one group?'' questions use the ``sharing'' model of division.  
\item ``How many groups?'' questions use what is sometimes called the ``measurement'' model of division.  
\end{itemize}
Division problems can have many types of answers, depending on the numbers and the context:
\begin{enumerate}
\item exact division (i.e., no remainder)
\item complete division (e.g., with a fractional part)
\item quotient and remainder (i.e., two numbers with different meanings)
\item just the quotient (i.e.,ignoring a non-zero remainder, rounding down)
\item quotient rounded up 
\item just the remainder
\end{enumerate}
\end{teachingnote}
Solve each of the problems below, explain your reasoning, and indicate whether the problem is asking ``\textbf{How many in one group?}'' or ``\textbf{How many groups?}'' or something else entirely.

\begin{problem}
There are a total of $35$ hard candies. If there are $5$ boxes with an
equal number of candies in each box---and all the candy is accounted
for, then how many candies are in each box? What if you had $39$
candies?
\end{problem}

\begin{problem}
There are a total of $28$ hard candies. If there are $4$ candies in
each box, how many boxes are there? What if you had $34$ candies?
\end{problem}


\begin{problem}
There is a total of 29 gallons of milk to be put in 6 containers.  If
each container holds the same amount of milk and all the milk is
accounted for, how much milk will each container hold?
\end{problem}

 
\begin{problem}
There is a total of 29 gallons of milk to be sold in containers holding
6 gallons each.  If all the milk is used, how many containers can be sold?
\end{problem}

\begin{problem}
There is a total of 29 gallons of milk to be sold in containers holding
6 gallons each.  If all the milk is used, how much milk cannot be sold?
\end{problem}
 
\begin{problem}
If there are 29 kids and each van holds 6 kids, how many vans do
we need for the field trip?
\end{problem}


%\begin{problem}
%Paolo has a total of $48$ outfits (shirts and pants) he can wear. If
%he has $8$ shirts, how many pants does he have?
%\end{problem}

%\begin{problem}
%A chart has $72$ cells and $8$ rows. How many columns does it have?
%\end{problem}

%\begin{problem} 
%A rectangle has a length of $6$ inches and an area of $42$ square
%inches. What is its width?
%\end{problem}

%\begin{problem} 
%Can you think of a division problem that is fundamentally different
%from the problems above?
%\end{problem}

%\begin{problem} 
%In the context of the problems above, what might ``division by zero''
%mean?
%\end{problem}

\end{document}
