%\documentclass[handout]{ximera}
\documentclass[nooutcomes]{ximera}

\input{../preamble}

\title{They'll Fall for Anything!}
\author{Bart Snapp and Brad Findell}

\outcome{Learning outcome goes here.}

\begin{document}
\begin{abstract}
  We explore some common misconceptions.
\end{abstract}
\maketitle

\label{A:fallForAnything}

What is incorrect about the following reasoning? Be specific!

\begin{problem}
Herman says that if you pick a United States citizen at random, the
probability of selecting a citizen from Indiana is because Indiana is
one of 50 equally likely states to be selected.
\end{problem}


\begin{problem}
Jerry has set up a game in which one wins a prize if he/she selects an
orange chip from a bag.  There are two bags to choose from.  One has 2
orange and 4 green chips.  The other bag has 7 orange and 7 green
chips.  Jerry argues that you have a better chance of winning by
drawing from the second bag because there are more orange chips in it.
\end{problem}

\begin{problem}
Gil the Gambler says that it is just as likely to flip 5 coins and get
exactly 3 heads as it is to flip 10 coins and get exactly 6 heads
because
\[
\frac{3}{5} = \frac{6}{10}
\]
\end{problem}

\begin{problem}
We draw 4 cards without replacement from a deck of 52.  Know-it-all
Ned says the probability of obtaining all four 7's is
$\frac{4}{\binom{52}{4}}$ because there are ways to select the
$\binom{52}{4}$ 4 cards and there are four 7's in the deck.
\end{problem} 

\begin{problem}
At a festival, Stealin' Stan gives Crazy Chris the choice of one of
three prizes---each of which was hidden behind a door.  One of the
doors has a fabulous prize behind it while the other two doors each
have a ``zonk'' (a free used tube of toothpaste, etc.).  Crazy Chris
chooses Door \#1.  Before opening that door, Stealin' Stan shows Chris
that hidden behind Door \#3 is a zonk and gives Chris the option to
keep Door \#1 or switch to Door \#2.  Chris says, ``Big deal.  It doesn't
help my chances of winning to switch or not switch.''
\end{problem}

\end{document}
