\documentclass[handout,nooutcomes, noauthor]{ximera}

\title{Probably More Probability}

\begin{document}
\begin{abstract}
\end{abstract}
\maketitle

\begin{problem}
The Diet-Lite restaurant offers 3 entr\'{e}es, 4 side dishes, 2 desserts, and 5 kinds of drinks.
\begin{enumerate}
    \item If you picked a dish randomly, what is the probability you would select a dessert?
    \item If you randomly selected a dinner with one entr\'{e}e, one side dish, one dessert, and one drink, what is the probability that your dinner would include orange juice as the beverage? (Assume only one of the drink choices is OJ.)
\end{enumerate}
\vfill
\end{problem}


\begin{problem}
Suppose an Ohio License plate consists of two letters followed by two digits followed by two letters. Also assume that license plates are distributed randomly, that no license plates are forbidden, and that repeats of letters or digits are allowed (some of these things are not true in life, but assuming them makes our calculations easier).
\begin{enumerate}
    \item What is the probability that you will get the license plate $AA00AA$?
    \item What is the probability that the first letter of your license plate will be $J$?
    \item What is the probability that your license plate will have the letter $Z$ somewhere on it?
\end{enumerate}
\vfill
\end{problem}

\newpage
\begin{problem}
A class has 10 children, three of which are named Jay.  
\begin{enumerate}
\item What is the probability that if you choose a group of three children, exactly one of them will be named Jay?
\item What is the probability that if you choose a group of three children, at least one of them will be named Jay?
\end{enumerate}
\vfill
\end{problem}





\end{document}