\newpage
\section{Shemp: A Prequel to Home Base}\label{A:shemp}

\begin{prob}
Long before recorded history, Shemp, a shepherd, wanted to make sure all his sheep returned from the field.  But back then, no one had any concept of ``counting.''  One day, Shemp found a big pile of necklaces big enough for sheep.  Shemp figured out a scheme, using a bunch of pegs hammered into his cave wall, that would help him know whether his sheep had come in from the field.  What do you think his scheme was?
\end{prob}

\begin{prob}
Eons later, two Stooges (shepherds who were using Shemp's necklace scheme) wanted to know who had ``more'' sheep.  They still had no concept of ``counting,'' but Moe figured out how to solve the problem by using the necklaces.  What do you think Moe did?
\end{prob}

\begin{prob}
Centuries later, the Stooges discovered that they could make marks on their cave walls instead of using pegs and necklaces.  So, for example, Grog might have this written on his cave wall:  $||||||||||||||||||||||||$ while Greg would have this on his cave wall:  $||||||||||||||||||$.  To see if their sheep had all returned, they would just point to one mark, then go on to the next, and the next, and so on (saying ``ni'' for each mark).  Now, when arguing over which had more, each could bring a papyrus copy of the cave wall markings.  How might this system work?
\end{prob}

\begin{prob}
Finally, Larry the Finer, invaded Stooge Valley and wanted to know ``how many'' sheep each shepherd had.  Since the Stooge People had never thought of the idea of ``how many,'' Larry told them to give different names to each of the marks on their wall.  How might this work?  
\end{prob}

\begin{prob}
But Larry was impatient.  If a shepherd's wall looked like $$|||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||||$$
he couldn't tell quickly how many sheep there were.   Describe two or three steps that would solve Larry's problem.   
\end{prob}

\begin{prob}
What does it mean for someone to be able to ``count''?
\end{prob}
