%\documentclass[handout]{ximera}
\documentclass[nooutcomes,handout]{ximera}

\input{../preamble}

\title{Constructible Numbers, Part 2}
\author{Bart Snapp and Brad Findell}

\outcome{Learning outcome goes here.}

\begin{document}
\begin{abstract}
  We continue to use algebra to understand compass and straightedge
  constructions.
\end{abstract}
\maketitle

\begin{problem}
Suppose you are given points $(h, k)$, and $(p, q)$ with integer coordinates?  
\begin{enumerate}
\item Write an equation of the circle with center $(h, k)$ and containing the point $(p, q)$.  
\item What arithmetic operations were involved in writing your equation of the circle?  
\item What can you conclude about the numbers that are coefficients in your equation?   
\end{enumerate}
\vfill
\end{problem}

\begin{teachingnote}
Finding the intersection of a circle and a line involves a substitution and then solving a quadratic equation:  rational operations and extracting square roots.  Finding the intersection of two circles involves first subtracting the two circle equations to remove all squared terms.  Then we have a line.  Our students might not have enough experience with these skills.  
\end{teachingnote}

\begin{problem}
Solve the following equations simultaneously
$$(x-3)^2+(y-2)^2 = 14$$
$$ y = x + 4$$
\vfill
\end{problem}

\newpage

\begin{problem}
Solve the following equations simultaneously
$$(x-3)^2+(y-2)^2 = 18$$
$$ y = x + 5$$
\vfill
\end{problem}

\begin{problem}
Solve the following equations simultaneously
$$(x-3)^2+(y-2)^2 = 12$$
$$ y = x + 4$$
\vfill
\end{problem}

\newpage

\begin{problem}
Solve the following equations simultaneously
$$(x-3)^2+(y+2)^2 = 4$$
$$(x-1)^2+(y-2)^2 = 9$$
\vfill
\end{problem}

\begin{problem}
Solve the following equations simultaneously
$$(x-3)^2+(y+2)^2 = 4$$
$$(x+1)^2+(y-2)^2 = 9$$
\vfill
\end{problem}

\newpage

\begin{problem}
Suppose you are given equations of the form 
$$x^2 + ax +y^2+by = c$$
$$x^2 + dx +y^2+ey = f$$
where $a$, $b$, $c$, $d$, $e$, and $f$ are all integers.  
\begin{enumerate}
\item What kind of geometric objects do these equations describe in the $xy$-plane?  
\item What arithmetic operations would you use to solve the equations simultaneously? 
\item What can you conclude about the numbers $x$ and $y$ that are the (simultaneous) solutions of these equations?  
\item How will your answers change if $a$, $b$, $c$, $d$, $e$, and $f$  are all rational numbers?  
\end{enumerate}
\end{problem}

\begin{problem}
Based on the previous problems, if you begin with a coordinate system with only integer coordinates, how would you describe the set of all numbers (coordinates) that are constructible via lines and circles?  
\vfill
\end{problem}

\begin{problem}
Considering that all compass and straightedge constructions are about lines, circles, and their intersections, what do your results about coordinate constructions imply about compass and straightedge constructions?  
\vfill
\end{problem}

\begin{problem}
Name some numbers that are \textbf{not constructible} with compass and straightedge. 
\vfill 
\end{problem}

\end{document}
