%\documentclass[handout]{ximera}
\documentclass[nooutcomes]{ximera}

\input{../preamble}

\title{Isosceles Bisectors}
\author{Bart Snapp and Brad Findell}

\outcome{Learning outcome goes here.}

\begin{document}
\begin{abstract}
  We think about isosceles triangles.
\end{abstract}
\maketitle

\begin{teachingnote}
Students typically draw a new line. They must choose which line they are drawing and then they must assume no additional properties.  Then they look for triangle congruence.  Median: SSS.  Altitude: HL.  Angle bisector: SAS. Perpendicular bisector: doesn't work because it might not contain the opposite vertex.  

Work through analogous ideas for the converse. 
\end{teachingnote}

\begin{theorem}[Isosceles Triangle Theorem]
If two sides of a triangle are congruent, then the angles opposite those sides are congruent. 
\end{theorem}

\begin{problem}
Prove the Isosceles Triangle Theorem.  (Hint: In a previous activity, you noticed that in most triangles the median, perpendicular bisector, angle bisector, and altitude to a side lie on four different lines.  So if you draw a new line in your diagram, be sure to decide which of these lines you are drawing.)
\vfill
\end{problem}

\begin{problem}
Use your proof to show that a median, perpendicular bisector, angle bisector, and altitude turn out to be the same line.
\vfill
\end{problem}

\newpage
\begin{problem}
Prove the Isosceles Triangle Theorem without drawing another line.  Hint:  Is there a way in which the triangle is congruent to itself? 
\vfill
\end{problem}

\begin{problem}
State the converse of the Isosceles Triangle Theorem and prove it.  
\vfill
\end{problem}

\newpage
\begin{problem}
Prove that the points on the perpendicular bisector of a segment are \emph{exactly those} that are equidistant from the endpoints of the segment.  Note that the phrase \emph{exactly those} requires that we prove a simpler statement as well as its converse:   
\begin{enumerate}
\item Prove that a point on the perpendicular bisector of a segment is equidistant from the endpoints of that segment.
\item Prove that a point that is equidistant from the endpoints of a segment lies on the perpendicular bisector of that segment.
\end{enumerate}
\vfill
\end{problem}

\begin{problem}
Prove that the perpendicular bisectors of a triangle are concurrent.  Hint:  Name the intersection of two of the perpendicular bisectors and then show that it must also lie on the third one.  (This is a standard approach for showing the concurrency of three lines.)  
\vfill
\end{problem}

\newpage
\begin{problem}
Draw a line (neither horizontal nor vertical) and a point not on the line.  Describe how to find the \emph{exact} distance from the point to the line. 
\vfill
\end{problem}

\begin{problem}
Prove that the points on an angle bisector are \emph{exactly those} that are equidistant from the sides of the angle. 
\vfill
\end{problem}

\begin{problem}
Prove that the angle bisectors of a triangle are concurrent. 
\vfill
\end{problem}
\end{document}
