\newpage
\activity\section{Big Primes}


All over the world, computers are churning away, trying to to find the
next prime. This is what is known as the Great Internet Mersenne Prime
Search---or (don't laugh) \index{GIMPS}GIMPS. If you are willing, you
too can play along and donate time on your computer. Every year or so
a new prime is found, and glory-everlasting\footnote{This might be a
slight exaggeration.} goes to the person who finds it! I will not
attempt to be up to date. Please use some inter-web-search-doo-hicky
or whatever to find the largest known prime. Please write the actual
number below:

\vspace{1in}

There is little doubt in my mind that this largest known prime will be
of the form:
\[
M_n = 2^n -1
\]
where $n$ is some number. Numbers of this form are
called \textit{Mersenne} numbers.\index{Mersenne numbers}

\begin{prob} 
Make a table of the Mersenne numbers $M_n$ for $n$ from $1$ to $10$. 
\end{prob}

\begin{prob} 
Which of the numbers in your table are prime? Make a conjecture.
\end{prob}

\begin{prob} 
Continue your table out to $M_{14}$. Check your conjecture up to this
point.
\end{prob}


\begin{prob} 
Compare this activity to the activity \textit{Whoops!} What do you
notice?
\end{prob}


By the way, we should add---$M_{n}$ grows very fast as $n$
grows. Let's do some comparisons:

\begin{prob} 
Which is bigger, the largest known Mersenne prime or the number of atoms
in \textit{you}?
\end{prob}

\begin{prob} 
Which is bigger, the largest known Mersenne prime or the number of
atoms in the planet Earth?
\end{prob}

\begin{prob} 
Which is bigger, the largest known Mersenne prime or the number of
atoms in the Sun?
\end{prob}

\begin{prob} 
Which is bigger, the largest known Mersenne prime or the number of
atoms in the visible Universe?
\end{prob}


