\documentclass[handout]{ximera}

\input{../../preamble}


\title{The Shepherd's Necklace}
\author{Vic Ferdinand, Betsy McNeal, Robin Pemantle, Jenny Sheldon}

\begin{document}
\begin{abstract}\end{abstract}
\maketitle




\begin{problem} 
A small shepherd boy was responsible for taking 23 sheep to pasture each day.  These sheep would graze freely all day, and in the evening the boy was to round them all up and drive them home.  Unfortunately, the boy could not count, so each day he had difficulty deciding whether he had found all of the sheep before returning home. 

\begin{enumerate}

\item	Devise a method (other than teaching the boy to count) that would make it possible for him to be sure he had all of the sheep.
% \vfill
% \item In this course, we will ask you to explain your work.  Let's begin by thinking about what an explanation for the previous question would look like.  Start by making an outline: what are the most important points you would want to talk about if you were explaining your answer to a child?  What details would you add?
% \vfill
% \item The most important part of your explanation is talking about why each step is true mathematically.  Go back to the outline you just wrote, and explain why each step is true.  Imagine you are talking to a skeptic!
% \vfill
% \item Now use your outline to write out your explanation.

\vfill


%\end{enumerate}
%\end{problem}
\vfill


MOVE ON ONLY WHEN INSTRUCTED!

\newpage
% \begin{problem} Let's return to the fields: to the same pasture, the same shepherd boy, and the same $23$ sheep.
% \begin{enumerate}
\item 	The sheep's owner came up with a simple scheme.  It involved making, for each sheep, a string necklace that would fit around the sheep's neck.  In the morning, the boy would take the necklaces and place them all around his own neck.  What should he do with them in the evening?
%\vskip 1in

\item \label{ShepherdsNecklaceC} 	The ``necklace idea" spread throughout the region and soon all the young people tending sheep went to the meadows with string necklaces around their necks. One day, an argument arose between a shepherd and a shepherdess as to which was tending the largest number of sheep. Resolving the argument was made difficult since neither could count. Can you describe a method by which they could resolve their argument while leaving their sheep grazing undisturbed?
%\vskip 1.5in

\item 	What is possible to determine about two sets if one does not know how to count?

\vskip 0.5in
\end{enumerate}
\end{problem}
\begin{problem} 
%What does it mean to ``know how to count''?  How does it relate to the sheep and necklaces?  How could we define the number ``seven" if we knew no other quantities?

You give Malika some blocks. You tell her,``I've given you some blocks. Can you give me the same number of pencils?" How can Malika know she has given you the same number of pencils, without knowing how to count? How would she know if she gave you too many? Too few? Use the definition of one-to-one correspondence to explain your reasoning.

\end{problem}

\begin{problem}
The Japanese counting words are (in order): ichi, ni, san, shi, go, roku, sichi, hachi, kyu. What is the Japanese word for 6? How do you know?

\end{problem}



\newpage
\begin{instructorNotes}

This activity brings up a situation in which a one-to-one correspondence is useful for comparing two sets.  We use this opportunity to define one-to-one correspondence, and show the importance and power of one-to-one correspondence in comparing two sets without the use of counting. 
This is also the conceptual foundation for quantifying a set because quantifying or counting is just a one-to-one correspondence between the (ordered) number words and the objects.  
The final question in the activity asks the students to try to define what it means to know how to count. We bring up this activity later in the course any time we find a one-to-one correspondence, to begin to give students a sense that mathematics has structures that come up repeatedly.


With the first question, we begin by eliciting a variety of ideas from the students about how to be sure all the sheep are accounted for.  Students usually come up with ideas like assigning a color or letter of the alphabet to each sheep and then checking the sheep off of a list as they return.  Elicit a variety of ideas from the students. Usually, they are all correct answers, and fall in various categories.  We point out that their ideas all have something in common - the idea of one-to-one correspondence between the sheep and something that is easier to track.  We note that order is not significant for this, that is, all sheep have returned if all colors, letters, or necklaces have been accounted for.

%%
% We will use this activity on the first day of class to introduce the kind of group work we want students to do, and the kinds of explanations we will expect from them during the semester.  Once students have fully engaged with this idea, we step back to think about how we would explain our ideas to someone else, and engage the class in a discussion about what makes a good mathematical explanation.  We don't take the time to actually write out the explanation on the board, but putting up an outline (especially one generated by a student or the class as a whole) should be a good model for students.  We don't expect to get beyond the first page on the first day of the class, after talking about the syllabus and course expectations.
%%

The situation in part c requires comparison of the necklaces in order to see who has more without disturbing the sheep.  So we can point out that counting is not needed in order to determine greater, less than, or equal. At this point, we usually stop to define the term ``one-to-one correspondence''.



The last two questions help us to discuss what it means to know how to count.  This is often a point where students begin to realize that they have forgotten how much they struggled to learn things as young children.  We like to focus students on how even though counting may seem ``easy'' to them, it is not an easy thing for children to learn.
In order to know a quantity, for instance ``seven'', without knowing any other quantities, you need a ``master set'' with seven objects in it to compare any other set (obviously, we're just working discretely here, not continuously - no Dedekind Cuts, please!).  ``Seven'' is the attribute that all the sets that match with the master set share (shaking out all other attributes in which they are different, such as size, type, color, etc. of the actual set members). Malika's problem helps us to point out this idea. The final problem helps us to distinguish between the English words we use to describe numbers and the idea of quantities themselves. We have a one-to-one correspondence between the quantities and the number-words we use to describe them (and the numerals we use to write them down!) 


Relating the meaning of counting to children's development of the meaning of numbers is not part of this activity, so we normally discuss it as a mini-lecture from the instructor, or ask students if they can contribute their own observations from working with children. For example, when children first are learning to count, they mimic the number words (a social, not mathematical, construct) that others use.  Then children try to point and count objects, but sometimes miss or repeat some objects in their point-and-say scheme.  This is once again an imitation of a social tradition and children often first believe that there are ``fewer'' objects in a clumped-up pile than in a spread out set.  Then finally the child ``realizes'' the importance of a one-to-one correspondence between the words in the socially accepted order and the objects.  Children then conclude that the number of objects is the same as the last counting word spoken, but they may have the order incorrect and even, sometimes, restate the last number word as the total even if one of the objects counted earlier is removed.


{\bf Suggested Timing:} We take about 15 minutes in small groups to discuss the activity (and encourage them to finish answering part (a) before turning the page).  
Discussing their ideas as a whole class usually takes about 20 minutes before summarizing the key points in counting.  



\end{instructorNotes}

\end{document}
