\documentclass{ximera}

\input{../preamble.tex}

\title{Multiplication with Integers}
\author{Vic Ferdinand, Betsy McNeal, Jenny Sheldon}

\begin{document}

\begin{abstract}
We look at multiplying integers.
\end{abstract}
\maketitle











\end{document}



Now, things get a little bit more complicated when dealing with a more complicated operation such as multiplication, where the “wholes” for each of the three quantities in a problem are different (# Groups   # Objects in each Group = # Objects).  So we need to be a little more careful in how we phrase our problems.
As always, let’s start with what we know to lead us to what we don’t know:  Let’s model (3) (5) with a problem that we’ll also want to use for (3) (-5), (-3) (5), and (-3) (-5).  Because multiplication is “repeated addition”, we want to try to stay close to what we’ve already accomplished with addition and subtraction.
For (3) (5), we’ll be joining 3 groups with 5 objects in each of those groups.  That is, it’s already “addition” from the start.  Since our addition (and missing addend subtraction) problems asked “what is our net worth now”, let’s stay with that, which forces our objects (our “answer”) to be dollars (positive means our final net worth is “ahead” and negative means we wind up “in debt”).  
Thus, for (3) (5), we’ll have 3 groups with 5 objects in each group, which is 3 groups with a positive 5 dollars in each group.  So, each group is going to be, in this example, a check for $5 (so a positive second number means we’re talking about checks.  A negative second number means we’re talking about bills for that amount).
The second issue is whether we’re receiving or sending these checks or bills (i.e., receiving or sending these objects).  That is, will we be receiving a bunch of checks or bills for the same amount or will we be sending those checks or bills for the same amount?  Because we have not dealt with the sign of the first number, this is what that will mean.  That is, (3) (5) will be “we receive 3 checks, each worth $5”.  The “receive” is from the first number (3) being positive.  The “checks” is from the second number (5) being positive.
Finally, our stories for addition always asked “what is our net worth now?”  However, our “starting” net worth was always our first number (e.g., for 4 + 5:  “We start ahead $4…”).  Here, we can’t do that for (3) (5) because 3 is telling us how many checks or bills we are receiving or sending.  Thus, if our final question is still going to be “what is our net worth now?”, what should we assume our “starting net worth” to be with each problem?
Thus, let’s put this all together:  The story problem for (3) (5) will be (fill in the blank and circle the appropriate choices:  
“We start with a net worth of ____.  We (receive or send) 3 (checks or bills) for $5 each.  What is our net worth now?”
Now do the same for each of (3) (-5), (-3) (5), and (-3) (-5).  Write the story, model it with chips, and model it on a number line (Direction determined by “good or bad news”).  Finally, write an arithmetic expression for each involving repeated addition or repeated subtraction (depending on whether we’re receiving or sending).

2.  Making sense by numerical patterns:
      a.  Starting with knowing (8)  (5) = 40 (and knowing that (positive)  (positive) = positive), work your way down one-by-one on the second number until you reach (8)  (0) = 0.  Notice a pattern in the answers and use that pattern to predict what (8)  (-1) should be.  Then predict what (8)  (-2) and (8)  (-3) should be.
      b.  Starting with knowing (8)  (-5) = -40 (and knowing that (positive)  (negative) = negative from part (a)), work your way down one-by-one on the first number until you reach (0)  (-5) = 0.  Notice a pattern in the answers and use that pattern to predict what (-1)  (-5) should be.  Then predict what (-2)  (-5) and (-3)  (-5) should be.

3.   Assuming that we know that the associative property of multiplication holds for all signed numbers AND that we know that unlike signs multiply to be a negative, what is wrong with the following argument?:
We want to find out what (-3)  (-5) should be (or in general, what a negative times negative should be).  
Step #1:  Well, (-3) = (-1)  (3).  
Step #2:  So, (-3)  (-5) should be   (-1)   (3)  (-5)   (i.e., rewriting (-3)).  
Because we know the associative property of multiplication and that a positive times negative = negative, we pair up the (3)  (-5) to get (-15).
Thus, we have (-1)   (3)  (-5) = (-1)   (-15) = 15.  
Thus, (-3)  (-5) = 15 (and, thus, by similar argument, negative times negative must be positive).

4.    Back in school, you probably learned that distance = rate x time.  In this exercise, we’ll look at it as distance = time x rate, so that the number of groups is the first factor.  For example, if you traveled for 3 hours at a rate of 50 miles per hour, you would have traveled a distance of 3 x 50 = 150 miles (objects are miles and groups are hours).  In other words, your position has shifted 150 miles from where you started.
       Let’s expand this idea to take advantage that we can speak of time in terms of time in the past (“ago”:  considered to be negative time) and time in the future (positive time).  We can also speak of position (as opposed to distance, which is always positive) in terms of west (negative) and east (positive).  
       In this scheme, we can rewrite our 3 x 50 problem in the following way:  Let’s assume we are currently at home, which sits on a west-east highway.  Let’s say we’ve been traveling eastward at 50 miles per hour for awhile and will continue to do so.  Where will we be 3 hours from now (in the future)?  That is, what is our position 3 hours from now (as opposed to asking how far did we travel)?
              In this scheme, what would the story problem (and answer) be for:
3 x (-50)?  
(-3) x 50?  
(-3) x (-50)?  

