\documentclass{ximera}
\input{../../preamble}
\title{Go Climb A Tree!}

\begin{document}
\begin{abstract}
\end{abstract}
\maketitle



\begin{problem}
\begin{enumerate}
\item Starting with the statement ``$\frac25$ of the class are girls'', write a story problem that is modeled by the expression $\frac37 \times \frac25$.  Once you have a story, solve it using pictures and remind yourself why multiplying fractions is the same as multiplying the numerators and the denominators.
\item Restate your story as a probability problem and use a tree diagram to illustrate the solution.
\end{enumerate}
\end{problem}

\begin{problem}\label{NBAStar}
Suppose an NBA superstar has been a 70\% free throw shooter throughout his long and storied career.  (That is, he has made 70\% of all free throw shots he has taken.)  Suppose he is now in a ``one-and-one'' situation: he shoots a free throw.  If he makes it, he gets one point and shoots one additional free throw, worth one point.  If he misses, he gets zero points and does not shoot another free throw.  What is the probability that:
\begin{enumerate}
\item he makes zero points?
\item he makes one point?
\item he makes two points?
\end{enumerate}
\end{problem}

\begin{problem}
Why does it make sense that the probabilities in Problem 2 add up to 1?
\end{problem}

\begin{problem}
The Reds and the Cardinals face each other in a playoff series.  The probability of the Reds winning any given game is 0.3, or 30\%.  Say that the two teams will play a best-of-three series, which means that the first team to win two games will win the whole series.

\begin{enumerate}
\item What is the probability that the Reds win games 1 and 3 to win the series?
\item What is the probability that the Reds win the series in exactly three games?
\end{enumerate}
\end{problem}

\begin{problem}
The Reds and the Yankees face each other in a playoff series (maybe the World Series??).  The probability of the Reds winning any given game is 0.3, or 30\%.  Say that the two teams will play a best-of-seven series, which means that the first team to win four games will win the whole series.

\begin{enumerate}
\item What is the probability that the Reds win games 1, 3, 4, and 7 to win the series?
\item What is the probability that the Reds win the series in exactly six games?
\end{enumerate}
\end{problem}


\newpage

\begin{instructorNotes}
This activity introduces students to the notion of using a tree diagram to help calculate probabilities, and emphasizes the relationship between fraction multiplication as we've seen it in 1125 and fraction multiplication in the situation of probability.  The first problem should be carefully considered so that the students make the right connections before getting started.  You may want to go through the second problem as a class, in order to help students see what they should do for later problems.

Make sure to emphasize the operations used here - why do we multiply?  Why do we add?

With a reduced counting section, the final part of the final problem may be too difficult for students unless they realize that they have counted this situation several times before - any time they have chosen a group of three things out of five (or computed $_5C_3$).  However, this may look like $_6C_4$ at first to students, which they have likely not calculated yet.
\end{instructorNotes}

\end{document}