\documentclass[nooutcomes,noauthor]{ximera}
\input{../../preamble}
\title{You Can Count On It Getting More Difficult}

\begin{document}
\begin{abstract}
\end{abstract}
\maketitle





\begin{problem}
You choose a card from a deck consisting of 10 distinct cards, and then you roll a six-sided die.  How many different outcomes are possible?
\end{problem}

\begin{problem}
You choose two cards at the same time from a deck consisting of 10 distinct cards, and then you draw three marbles at the same time from a bag consisting of five different marbles.  How many different outcomes are possible?
\begin{enumerate}
\item Before you solve this problem, list at least three potential outcomes.
\item Have you seen any parts of this problem before?
\end{enumerate}
\end{problem}

\begin{problem}
A class has 10 children, three of which are named Jay.  How many ways are there to choose a group of three children containing at least one person named Jay?
\end{problem}


\newpage

\begin{instructorNotes}
This activity consists of more complicated counting problems.  It is probably impossible to finish all of them in the time provided!

Students should, at this point, be comfortable with comparing and contrasting problems they are trying to solve with problems they have already solved.  This activity continues to emphasize this practice.

This activity also introduces counting two separate events using multiplication.  You should be careful to review the groups and objects per group for the multiplication.
\end{instructorNotes}


\end{document}