\documentclass[noauthor, nooutcomes]{ximera}

\input{../preamble}


\title{Work on wording}
\author{Vic Ferdinand, Betsy McNeal, Jenny Sheldon}

\begin{document}
\begin{abstract} 
In this activity, we try to understand why we can use ``one tenth'' to represent two different ideas.
\end{abstract}
\maketitle



\begin{problem}
\begin{enumerate}
	\item Use bundling to explain what we mean when we say ``one tenth" and we mean $0.1$. Draw a picture to illustrate your thinking.
	\item Use the meaning of fractions to explain what we mean when we say ``one tenth'' and we mean $\frac{1}{10}$. Draw a picture to illustrate your thinking.
	\item Use your work in the previous two parts to explain why it makes sense to use the same words for both of these ideas. Why is $0.1 = \frac{1}{10}$ when they have different meanings?
\end{enumerate}

\end{problem}


\begin{problem}
Use equivalent fractions and your work in the previous problem to convert $\frac{1}{25}$ into a decimal. For a challenge, try $\frac{11}{40}$ as well! Why does your method make sense?
\end{problem}

\begin{problem}
Give three examples of fractions you could use the same process as the previous problem to find their decimal equivalent, and one example of a fraction that you could {\bf not} use the same process. Explain which is which and how you know.
\end{problem}



\begin{problem}
Use the meaning of fractions to shade $\frac{3}{8}$ of the grid below. (Do not begin with the decimal for $\frac{3}{8}$!) What does your shading tell you that the decimal representation for $\frac{3}{8}$ should be? Use another method to confirm that you got the correct answer.
\begin{center}
    \begin{tikzpicture}[scale=0.5]
    \draw [step=1, help lines] (0,0) grid (10,10);
    \end{tikzpicture}
\end{center}
\end{problem}


\begin{problem}
Use the meaning of fractions to shade $\frac{2}{3}$ of the grid below. What does your shading tell you that the decimal representation for $\frac{2}{3}$ should be? Use another method to confirm that you got the correct answer.
\begin{center}
    \begin{tikzpicture}[scale=0.5]
    \draw [step=1, help lines] (0,0) grid (10,10);
    \end{tikzpicture}
\end{center}
\end{problem}



\begin{problem}
So far, we've been changing fractions into decimals. Let's go the other way! Use what you know about the base-ten system to write the following decimals as fractions.
\begin{enumerate}
    \item $0.04$
    \item $0.483$
    \item $0.0003205$
\end{enumerate}
\end{problem}



 \newpage
\begin{instructorNotes}

{\bf Main Goal:} We would like students to be able to explain why $0.1$ is equal to $\frac{1}{10}$ (ie the first problem is the main idea here).

{\bf Overall Picture:} This is an example of an activity where students may first need to be convinced that these two ideas are different before we convince them that they are actually the same. Be ready to be patient!

 \begin{itemize}
 \item The first problem is designed to help students bring together the fractions and bundling.  The main goal is to see that $0.1$ is made when we take our unit and divide it into 10 equal parts, selecting one of those parts. But this is exactly our definition of $\frac{1}{10}$.
 \item Once we see that we can understand fractions with powers of 10 in the denominator as the same as decimals, our goal is to convert fractions to having such a denominator. These two examples (and the reverse problem later in the activity where we convert from decimals to fractions) should emphasize this point.
 
 \item We also hope that students see that the goal when changing to a decimal is to make a denominator that's a power of $10$. We don't have to express it that way (it's okay to say $10$, $100$, $1000$, and so on) but we want to recognize this as the goal.
  
  \item The Hundredths Grid (shading problems) are a very nice place to encourage students to be creative, and to see lots of different types of solutions presented. Here are some of the strategies students have used in the past, with the example of shading $\frac16$ of the grid (note that the actual problems are different fractions). Some students will cut the box into six equal pieces and shade one of them. Other students will shade 1 out of every 6 boxes (connecting the $\frac{1}{6}$ to the ratio $1:6$). Still other clever methods have been seen in the past! This is a nice place to remind the students of the relationship between a fraction, a rate, and a ratio.


\item The final question gives us a chance to talk briefly about irrational numbers. Depending on time, you can have the students come up with examples of irrationals. The most common first examples are famous irrationals like $\pi$, $e$, $\sqrt{2}$. You can also encourage students to come up with a ``homemade'' irrational number such as $0.01001000100001\dots$, which encourages students to think about patterns that are not repeating, and perhaps also helps students to understand that there are a lot more irrational numbers out there than they perhaps imagined! 


 \end{itemize}


{\bf Suggested Timing:}  Start by giving students 5-10 minutes to think about the first three problems, then spend about 20 minutes in discussion with groups presenting their work. Next, give students about 10 minutes to think about the rest of the worksheet, and then discuss as much as you have time for.
\end{instructorNotes}





\end{document}
