\documentclass[nooutcomes, noauthor, handout]{ximera}
\input{../../preamble}
\title{Symmetry}

\begin{document}
\begin{abstract}

\end{abstract}
\maketitle

We say an object has symmetry if we apply some transformation to that object, and the result looks exactly the same as before we applied that transformation.  Let's think about what this definition really means.



\begin{problem}
This problem is brought to you by the letter H.
\begin{image}
\begin{tikzpicture}
\draw[ultra thick] (0,0)--(0,4);
\draw[ultra thick] (3,0)--(3,4);
\draw[ultra thick] (0,2)--(3,2);
\end{tikzpicture}
\end{image}
Find all of the symmetries of the letter H.  In other words: apply a transformation to the letter H.  Does it still look like H?  Is it still in the same place on the page?  Are both of these questions important?

\end{problem}



\begin{problem}
Draw a rectangle and then show all of the symmetries of your rectangle. How many different rotation symmetries are there? How many reflection symmetries?
\end{problem}



\begin{problem}
Use your work in the previous problem to explain why the opposite sides of a rectangle are equal.
\end{problem}




\begin{problem}
Draw an isosceles triangle which is not equilateral and then show all of the symmetries of your triangle. How many different rotation symmetries are there? How many reflection symmetries?
\end{problem}


\begin{problem}
Use your work in the previous problem to explain why the base angles of an isosceles triangle are equal.
\end{problem}


\begin{problem}
What kind of symmetries does a circle have? Explain your thinking, drawing pictures to illustrate your ideas.
\end{problem}



\begin{problem}
Draw a design that has a rotation symmetry but no reflection symmetry. Explain how you know your answer is correct.
\end{problem}


\begin{problem}
What would it mean for a design to have translation symmetry? Draw examples to illustrate your thinking.
\end{problem}






\newpage
\begin{instructorNotes}

{\bf Main goal:} We introduce the notion of symmetry and produce examples.

{\bf Overall picture:} Tracing paper can be a huge help for this activity - encourage students to make two copies of their letters, shapes, or designs to experiment with.

\begin{itemize}
	\item The main thing we want to accomplish here is practicing with our definitions of rotations, reflections, and translations. In discussion, be sure to have students specify their transformations exactly and review their definitions as needed.
	\item A secondary goal here is to see that symmetries tell us about properties of shapes in a different way than the triangle congruence theorems we already studied. If students relate this work to our triangle congruence work, highlight the ideas of transformations underlying both congruence and symmetry.
	\item You should discuss that when specifying a symmetry, one should make sure to specify the transformation exactly.  For instance, with a reflection symmetry, students should give the line of reflection.  With a rotational symmetry, students should give the angle of rotation.
	\item With rotation symmetry, you want to bring up the idea of the order of the rotation, as in the text.
	%\item In discussion, you want to work through at least four different letters including the H.
	
	\item The idea of translation symmetry may have come up as students tested the H, but we definitely want to discuss it with the final problem. Hopefully a student will come up with the idea of the object extending infinitely, or perhaps living on a sphere so that it has no ends. Students can struggle with translation symmetry if they don't understand that the design must end up in the same location on the page. 
	\item Discuss the case of the rectangle's symmetries carefully. You should have some students discuss using tracing paper or two copies of the shape and some students discussing the transformations that would produce the symmetries. It would also be great to discuss why the diagonals of a rectangle are not lines of symmetry (especially if a group discovers this for themselves).
	\item Using symmetry to show that the opposite sides of a rectangle are congruent uses the idea that transformations don't stretch or shrink figures - they preserve distances. So when we can match up the sides with a reflection (for example), we know they are exactly the same length as well. The problems about the isosceles triangle emphasize this connection again.
	\item The circle problem brings up many interesting issues, but one that you might discuss if you haven't before is the idea that a circle is the limit of regular polygons as we keep adding more and more sides. You should avoid the limit language unless your students are comfortable with this, but describing the idea can be helpful.
	\item If you get to student-created examples, try to present as many as you can.
\end{itemize}


{\bf Suggested timing:} Give students about 5 minutes to think about the letter H and then discuss to make sure everyone is on the same page about the definitions of symmetry. Then give students about 10 minutes to work on the rest of the page and spend the remaining time in discussion using as many examples as possible. Problems 1-3 illustrate the main points, problems 4-5 are great extensions, and farther than that is a bonus!
\end{instructorNotes}

\end{document}





