\documentclass{ximera}

\input{../preamble}

\title{Integers and Subtraction}
\author{Vic Ferdinand, Betsy McNeal, Jenny Sheldon}

\begin{document}
\begin{abstract}
    Let's try subtracting some integers!
\end{abstract}
\maketitle



Here are some subtraction problems we'll be working with throughout the activity.

\begin{enumerate}
    \item $5 - 2$
    \item $15 - (-3)$
    \item $(-4) - 9$
    \item $(-6) - (-12)$
\end{enumerate}

\begin{problem}
We had several different interpretations of subtraction.  Using our real-life integer contexts, write a comparison story and either a change or a put together story for $5 - 2$.

\end{problem}

\begin{problem}
Use our real-life integer contexts to write story problems for the remaining three problems.  You can choose the type of story, but you might find some easier than others!  Draw pictures explaining what a child just learning this concept might do to solve the problem.


\end{problem}

\begin{problem}
For each of the subtraction problems above, how would you use a number line to model the solution?  Draw and explain clearly, especially when subtracting a negative number.  Does your number line model any of the stories you wrote above?
\end{problem}

\begin{problem}
Write a story problem for each of the subtraction problems above which uses our checks-and-bills scenario.  Make sure to read your story over several times.  Does the sensible answer to your story match the answer to the subtraction problem?
\end{problem}


\newpage
\begin{instructorNotes}
This activity is the second of three that investigate operations with negative numbers.  The first is ``Where Have the Integers Been?'' and the third is ``Our Problems are Multiplying''.  There is an accompanying reading (\url{https://ximera.osu.edu/m4t/elementaryTeachersOne}) on integers that lays out some of the more complicated framework, and the instructor notes for ``Where Have the Integers Been?'' contain background information about this unit.

Our overall goal in this second activity is to make significant progress in the understanding of the checks and bills model, particularly as it pertains to addition and subtraction problems. Since this is the last problem, you may want to direct students to it early during class, or simply write the four problems on the board and direct students to write ``checks-and-bills'' stories. If students finish this early, they can return to some of the other ideas suggested in the activity.

This activity's questions are intentionally similar to those on the other handouts about integers, so that we can quickly remind our students of the previous activity, and then select the problems where they need the most practice.  We do not often discuss all of the problems on this page.

\begin{itemize}
    \item The first two problems can be used to bring up some of the difficulties of integer models.  Students can quickly find that their models don't make sense or aren't flexible enough when we begin to operate on integers.
    \item The third problem gives space for talking about number lines.  In this activity, we focus on how we can use the number line to see that ``subtracting a negative'' gives the same result as ``adding a positive''.  We try to focus as always on reasoning, and why the rule makes sense.
    \item The fourth problem gives space for working with checks and bills.  Again, we try to help students see that we can reason about ``subtracting a negative'', not just memorize a rule.  The checks and bills stories, properly done, gives credence to the ''rules'' developed in \#3 with the number line. 
    \item Students generally need quite a bit of practice with writing checks and bills stories that are both stories for the operation at hand as well as asking the proper question in their story (not giving away the sign of the answer in their question, or even asking a question in the first place).  This will be even more complicated with multiplication, so we try to get a good start here with subtraction (and addition if we didn't have much time for that on the previous day) is a good idea.
    \item There are a few different ways to write the checks and bills stories. One option is to begin with a given net worth and send the second number's worth away.  We then ask what the resultant net worth will be (if we send a negative amount, we are sending a bill that we assume will later be paid to us:  A good way to explain why subtracting a negative is equivalent to adding a positive!). A second option is to start with a given net worth and receive something in the mail to result in the given goal amount.  It's vital for students not to give away what was received (a check or a bill) so that the question can be asked of what was received (so the answer must include both amount AND sign). Both of these stories are change type, happening over time. We could also write a comparison story, but these aren't outlined in the online reading and so will require more creativity from students.
    \item Another item we pay attention to with these problems is which type of story students are writing.  As they are working with their stories and number lines, we continually ask, ``which type of story is this?'' to help students gain experience with identifying the structure.
    \item  If you use chips, a nice benefit is that when students need to take away an amount (or color) of chips that they do not start with, they must rename that starting amount so that it includes what must be physically taken away.  This should not be surprising because students have already experienced renaming in two other subtraction algorithms (place value and fraction).
\end{itemize}


{\bf Suggested Timing:} This activity takes us an entire class period.  We often do a 5-minute reminder of the previous work, then give students 5 minutes on number lines, and then 15 minutes to discuss the number line context.  We then give 5-10 minutes for checks and bills, and the remainder of the time to discussion.  We often swap the order of these contexts as well.
\end{instructorNotes}







\end{document}
