%\documentclass[handout]{ximera}
\documentclass[nooutcomes,noauthor]{ximera}
\input{../../preamble}

\title{I'm the Biggest Kid!}

\begin{document}
\begin{abstract}\end{abstract}
\maketitle



\begin{problem}
    The students in Ms. Smith's class come in from recess having an argument.  Ms. Smith asks the children about the argument, and it soon becomes clear that the children can't decide who is the biggest in the class.  Why might the children be confused about this question?
    
%    \begin{solution}
%    The term ``biggest'' isn't very descriptive or very precise.  When you are measuring the attributes of the children, one child may be taller than another, but may also be skinnier.
%    \end{solution}
   
\end{problem}


\begin{problem}
    To settle the matter, Ms. Smith divides her class up into groups, and gives each group a different item to use:
    \begin{itemize}
        \item One group gets a ball of yarn.
        \item One group gets a stack of math textbooks.
        \item One group gets an unsharpened pencil.
        \item One group gets a playground ball.
        \item One group gets a yard stick (a 3-foot ruler).
    \end{itemize}
    
    How could the children in each group use this item to help them answer the question?  Be extremely specific.
    
%    \begin{solution}
%    Answers should be creative!  Find as many attributes to measure as possible.
%    \end{solution}

\end{problem}

\begin{problem}
    Which item given to the groups is the most useful?  Which item is the least useful?
    
    Come up with at least one way that each item could be misused for answering this question.
    
%    \begin{solution}
%    Answers will depend on personal preferences, and which attributes you are measuring.  Notice that a ruler is useful for some things, but not useful for others!
%    \end{solution}
    
\end{problem}

\begin{problem}
    In what other ways could the children in Ms. Smith's class try to determine which child is the biggest?
    
%    \begin{solution}
%        Again, be creative!  One obvious attribute we cannot measure with the given tools is the weight of each child (for good reason)!  So, if we had a scale, we could measure which child has the largest weight.
%    \end{solution}
   
\end{problem}


\newpage

\begin{instructorNotes}

{\bf Main goal:} We introduce some of the concepts of measurement and get students reasoning about non-standard units. This activity is a success if the students come away with the idea that they need to choose an appropriate unit before they measure.

{\bf Overall picture:}

This activity is intended to introduce the subject of measurement.  The activity should bring up the need for measurement, the ambiguities that can occur when comparing different types of measurements, as well as give a gentle introduction to doing measurement.  We also encounter the idea of measuring different aspects of the same object, and how these different measurements can be somewhat independent.  Also included is some time to discuss children's misconceptions about measurement.

First problem:
\begin{itemize}
	\item         This question brings up the concept that the ``biggest'' is an ambiguous term.  Are the children debating who is the tallest?  Who weighs the most?  Whose arms are longest?  Have your class come up with as many ways as possible that a child could be considered the ``biggest'' - including some ways that seem a little unreasonable!
\end{itemize}


Second problem:
\begin{itemize}
 \item Here students are asked to begin to discuss measuring various aspects of the children in the class.  Most of the natural answers should be about 1-dimensional aspects of the children in the class, like their height or perhaps the distance around their middle.  Students should identify both the aspect they are measuring, as well as the unit they will use to measure (in that order).  Not all measurements need to have a unit (for instance, if the children will cut yarn to the height of each child, and then compare the lengths), but most will.  
        
\item         This is a good time to have a discussion about what a unit is, how we use it, and why choosing one is important!

\item Some of the students' suggestions will be a direct comparison between students, rather than a measurement. For example, we could cut the yarn into segments as tall as each child, and then line up the segments to see which is longest. This is a precursor to measurement, and thus an important idea! All measurement is a type of comparison, but we usually compare to a chosen unit so that we don't have to compare between objects.
        
\item         To use some of the unconventional items, students will have to be a little creative, and perhaps measure aspects of the children that are not one-dimensional.  For example, the children could try to measure the ``biggest'' child by how many math books they can hold up at once.  Another idea for the math books would be to try to figure out how many math books it would take to cover an outline of the child (the area of a child's silhouette).  The playground ball could be used to approximate the volume of a child.
        
 \item        During discussion, when students give a way to use one of the items, ask them to try to describe what, exactly, they are measuring.  Do they think that it's a length?  An area?  A volume?  None of the above?  This is also a good place to have students practice precision in their descriptions.  Make sure the entire class understands how the item is being used in measurement.

\end{itemize}

Third problem:
\begin{itemize}
       \item  The first question is a matter of opinion, but should be fun to discuss.  
        
      \item   The second question is far more interesting, in that students are asked to identify potential misconceptions that children may have about measurement.  For instance, when measuring with the pencil, the children could leave space between successive pencils, resulting in a measurement that is ``too short'' as well as inaccurate.  
        
      \item  Other common ``ruler misconceptions'' can also be discussed, here.  Sometimes a child will line up one end of an item to be measured with the ``1'' on the ruler, instead of the end of the ruler.  Another misconception is to put the ruler next to the object, and just read whichever marking is at the end, regardless of where the object starts.
\end{itemize}

Fourth problem:
\begin{itemize}
        \item This is another opportunity for students to be creative.  What items would they as teachers put on Ms. Smith's list?  Some measurement could also be done simply by comparison: for instance, we could just line up the children by height to see who is the tallest.  Discuss the pros and cons of the students' methods for determining who is the biggest kid in the class.
\end{itemize}




Take-aways from this activity: the concept of ``biggest" doesn't have much meaning unless we are more specific.  Draw out, but not yet answer or address many issues in measurement.  For example, what is an appropriate unit to use when measuring?  These ideas will be addressed in future activities - it's okay to not feel ``finished" with the discussion here.

{\bf Good language:} The more specific students can be with their ideas, the better. If they can make drawings or sketches of their ideas, this can help them communicate. Since we haven't yet talked about dimension, we don't expect this kind of reasoning from the students, but getting them to be specific about exactly what they are using as their unit will help us later on. For instance, if they are using the diameter of the playground ball as a length unit, we don't want them to say only ``I used the ball" but rather ``I used the distance all the way across the ball'' or something similar.

{\bf Suggested timing:} This activity and its discussion should take the entire class period.  Give students about 5 minutes to discuss the situation in Problem 1 amongst themselves, and then discuss with the class for about 5-10 minutes.  Let students think about problems 2-4 for the next 20 minutes, and use the rest of the period to discuss their ideas.

\end{instructorNotes}



\end{document}