\PassOptionsToPackage{numbers}{natbib} % passing an option to  natbib before tufte-book
\documentclass[justified,openany,nofonts]{tufte-book}


\bibliographystyle{amsplain}
\usepackage{teacherEd}


%\usepackage{microtype}

%\usepackage{lineno}
%\linenumbers


\usepackage{framed}
\usepackage{comment}
%\RequirePackage{xcolor,colortbl}% http://ctan.org/pkg/xcolor.  % Needed for secondDifferences activity. 

%\newenvironment{teachingnote}{\begin{framed}\begin{itshape}\begin{bfseries}\noindent Teaching Note:}{\end{bfseries}\end{itshape}\end{framed}}

\newenvironment{teachingnote}{\begin{framed}\noindent \textbf{Teaching Note:}}{\end{framed}}
\newcommand{\teachingnotes}{With Teaching Notes}
%\excludecomment{teachingnote}
%\newcommand{\teachingnotes}{}

\def\fixnote#1{\begin{framed}{\color{red}Fixnote: #1}\end{framed}}  % Allows insertion of red notes about needed edits
%\def\fixnote#1{}

%%% This sets how the enumerate command works
\renewcommand{\theenumi}{$(\mathrm{\arabic{enumi}})$}
\renewcommand{\labelenumi}{\theenumi}


\title{Numbers \\ and Algebra}
\author{\teachingnotes}
\publisher{Math 1165: Autumn 2018 \\ This document was typeset on \today.}

\begin{document}

\def\document#1{} %% Needed to add standards
\maketitle
%%%% Front matter
\input{../chapters/front}
%%%%

\vspace{1in}

\begin{teachingnote}
Activities to be added or revised:  
\begin{itemize}
\item Beckmann 1G
\item Product Puzzle from Connected Mathematics Project 
\item A.13. There are many factors to consider.  Needs revision. 
\item A.20. Picture Yourself Dividing.  Include $1\frac{3}{4 }\div \frac{1}{2}$. 
\item A.26. Poor Old Horatio.  Need new ratio table problems.
\item A.54  Second Differences.  Current version won't compile within notes.  
\end{itemize}
\end{teachingnote}

% Fixnote summary and other needs
%
% Highlight Deep Insights, such as
%  * The point-slope form of a line turns a non-proportional situation into a proportional one. 
%  * The sequence of partial sums of an arithmetic series is a quadratic function. 
%  * With common denominators, division of fractions ? 
%
% Fix the footnotes
%
%More problems needed: 
% 1.1. Other bases
% 1.2. Doubling and halving algorithm for multiplication 
%         Mysterious base two game
% 2.1. Models of integer arithmetic
% 3.1. Easy problems on ratio and proportional reasoning
% 	Ratio tables, going through 1, tables that aren?t ratio tables
% 3.2. More problem on quadratics, exponentials?  (Some in LaTeX comments.)
%
% Discussion:  Meaning of the equals sign somewhere
%


\newpage
%\pagenumbering{arabic}
%\pagestyle{fancy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% Sections to be included %%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{secnumdepth}{2}% turn on numbering for parts and chapters

\input{../chapters/algebra}

\input{../chapters/number}

\input{../chapters/ratio}

\input{../chapters/solve}

\input{../chapters/circle}


\appendix

\renewcommand{\theenumi}{$(\mathrm{\alph{enumi}})$}
\renewcommand{\labelenumi}{\theenumi}
\chapter{Activities}
%\addtocontents{toc}{\protect\setcounter{tocdepth}{0}}


%\input{../chapters/appendix/shemp.tex} 

\input{../chapters/appendix/bases.tex}  % Shelby and Scotty

\input{../chapters/appendix/hieroArith.tex}

\input{../chapters/appendix/blocksAlgo.tex}

\input{../chapters/appendix/blocksMult.tex}

\input{../chapters/appendix/compare.tex}

\input{../chapters/appendix/integerAddition}

\input{../chapters/appendix/integerMultiplication}

\input{../chapters/appendix/divisionMeaning.tex}

\input{../chapters/appendix/divisibility}

\input{../chapters/appendix/hallOfShoes.tex}

\input{../chapters/appendix/sieve.tex}

\input{../chapters/appendix/infPrimes.tex}

\input{../chapters/appendix/countingFactors.tex}

\input{../chapters/appendix/whyDoesItWork.tex}

\input{../chapters/appendix/prome.tex}

\input{../chapters/appendix/equivFrac.tex}

\input{../chapters/appendix/fracPictMean.tex}

\input{../chapters/appendix/fractionMultiplication}

\input{../chapters/appendix/divFractions.tex}

\input{../chapters/appendix/pictDivision.tex}

\input{../chapters/appendix/crossMult.tex}

\input{../chapters/appendix/hundredthsGrid}

\input{../chapters/appendix/repeat.tex}

\input{../chapters/appendix/decimals.tex}

\input{../chapters/appendix/ratioLaunch.tex}

\input{../chapters/appendix/ratios.tex}

\input{../chapters/appendix/ratioAddition.tex}

%\input{../chapters/appendix/solved.tex}

\input{../chapters/appendix/triathlete.tex}

\input{../chapters/appendix/dreadedStoryProblem.tex}

\input{../chapters/appendix/walkTheLine.tex}

\input{../chapters/appendix/constantAmount.tex}

\input{../chapters/appendix/constantRatio.tex}

%\input{../chapters/appendix/exponentRules.tex}  Replaced by Meanings of Exponents

\input{../chapters/appendix/meaningsOfExponents}

%\input{../chapters/appendix/doesntAddUp.tex}  Replaced by arithmetic series

%\input{../chapters/appendix/worldSeries.tex}  Moved to problems

% \input{../chapters/appendix/gertrude.tex}   % replaced by ConstantAmount and WorldSeries

% \input{../chapters/appendix/billy.tex}    %  replaced by ConstantRatio and WorldSeries

\input{../chapters/appendix/arithmeticSeries}

\input{../chapters/appendix/geometicSeries}

%\input{../chapters/appendix/secondDifferences}   % Requires \RequirePackage{xcolor,colortbl}, but then some of the Euclidean Algorithm sections won't compile.  

\input{../chapters/appendix/hieroAlg.tex}

\input{../chapters/appendix/otherSide.tex}

\input{../chapters/appendix/solvingQuadratics}

\input{../chapters/appendix/completeSquares}

\input{../chapters/appendix/vertex.tex}

%\input{../chapters/appendix/leastSquares.tex}

\input{../chapters/appendix/solvingCubics}

%\input{../chapters/appendix/otherCurves.tex}

\input{../chapters/appendix/sketchRoots.tex}

\input{../chapters/appendix/complexAddition.tex}

\input{../chapters/appendix/complexMultiplication.tex}

%\input{../chapters/appendix/deg2Ext.tex}

%\input{../chapters/appendix/inverseFunctions.tex}

%\input{../chapters/appendix/gaussianInt.tex}

%\input{../chapters/appendix/orderMod.tex}

%\input{../chapters/appendix/closeButNoCigar.tex}

\input{../chapters/appendix/traffic.tex}

\input{../chapters/appendix/factOrFiction.tex}

%\input{../chapters/appendix/pyramid.tex}

\input{../chapters/appendix/countOnIt.tex}

\input{../chapters/appendix/whichRoad.tex}

\input{../chapters/appendix/lumpyAndEddie.tex}

\input{../chapters/appendix/climbTree.tex}

\input{../chapters/appendix/fallForAnything.tex}


%\input{../chapters/appendix/harmtri.tex}

%\input{../chapters/appendix/whoops.tex}

%\input{../chapters/appendix/largePrime.tex}


%% \chapter{Enrichment Topics}

%% \input{../chapters/enrichment/continuedFractions.tex}

%% \input{../chapters/enrichment/music.tex}


%%%% Index and References
\input{../chapters/back}
%%%%

\end{document}
